\bfseries Equivariant convolution-based functions on tensor fields

### 2022-07-28 paper/main.tex

% typesetting adjustments
\makeatletter
\renewcommand\section{\@startsection {section}{1}{\z@}%
  {-3.25ex \@plus -1ex \@minus -.2ex}%
  {1.5ex \@plus .2ex}%
  {\raggedright\normalfont\large\bfseries}}
\makeatother
\setlength{\textwidth}{5.00in}
\setlength{\oddsidemargin}{3.25in}
\addtolength{\oddsidemargin}{-0.5\textwidth}
\setlength{\textheight}{9.40in}
\setlength{\topmargin}{-0.50in}
\setlength{\footskip}{0.25in}
\pagestyle{myheadings}
\markboth{foo}{Hogg \& Villar /Geometric filters for geometric images}
\newcommand{\figurerule}{\rule[1ex]{\textwidth}{0.2pt}}
\linespread{1.08}
\frenchspacing\sloppy\sloppypar\raggedbottom

\title{\bfseries%
Equivariant convolution-based functions
on tensor fields}
\author{Hogg, Villar, Blum-Smith, others}
\date{2022 June}

@@

\begin{definition}[$O(d)$ action on vectors and $k$-tensors] \label{def.tensors}
Let $O(d)$ act on vectors in $\mathbb R^d$ via the canonical action.
It also acts on rank-1 $k$-tensors in $(\mathbb R^d)^{\otimes k}$ by $g\cdot (v_{1}\otimes\ldots \otimes v_k) = (g\cdot v_1)\otimes \ldots \otimes (g\cdot v_k)$ and extending linearly to higher-rank tensors made from sums of rank-1 tensors.

It also acts on lists of $k$-tensors $((\mathbb R^d)^{\otimes k})^n$ by acting on each of the $n$ $k$-tensors by the same action.
%Let $W$ (aka ``the world'') be the comprehensive set of all vectors and tensors that exist.
%Note that $O(d)$ can be seen as acting on the world.
\end{definition}

@@

\begin{definition}[scalars and pseudoscalars]
We say that $a\in \mathbb R$ is a $0$-tensor or scalar iff $a$ is an invariant of the action.
Namely, $a(W) = a(g\cdot W)$ for all $g\in O(d)$.
We say that
$a\in \mathbb R$ is a $0$-pseudotensor or pseudoscalar iff $a(g \cdot W) = \det(g)\,a(W)$ for all $g\in O(d)$,
where $\det(g)$ is the determinant of the matrix representation $M(g)$ of $g$.
\end{definition}

\begin{definition}[pseudovectors and $k$-pseudotensors]
We say that $T\in (\mathbb R^d)^{\otimes k}$ if a $k$-pseudotensor if it is acted by $O(d)$ by the following action: $g\odot T = \det(g)\, g \cdot T$ where $g\cdot T$ is the cannonical action on tensors as in Definition \ref{def.tensors}.
\end{definition}

@@

\begin{definition}[permutations of tensor indices]
MSV WRITE THIS.
\end{definition}

@@

\section{Construction of equivariant geometric convolution filters}

In the above we defined nonlinear equivariant geometric operators acting on $d$-dimensional images on the $d$-torus.
These, in turn, are constructed from products of the outputs of equivariant geometric convolution filters.
Here we show how to find all such filters by a group-averaging technique.

>>>>

% typesetting adjustments
\makeatletter
\renewcommand\section{\@startsection {section}{1}{\z@}%
  {-3.25ex \@plus -1ex \@minus -.2ex}%
  {1.5ex \@plus .2ex}%
  {\raggedright\normalfont\large\bfseries}}
\makeatother
\setlength{\textwidth}{5.00in}
\setlength{\oddsidemargin}{3.25in}
\addtolength{\oddsidemargin}{-0.5\textwidth}
\setlength{\textheight}{9.40in}
\setlength{\topmargin}{-0.50in}
\setlength{\footskip}{0.25in}
\pagestyle{myheadings}
\markboth{foo}{Hogg, Blum-Smith, \& Villar /Geometric filters for geometric images}
\newcommand{\figurerule}{\rule[1ex]{\textwidth}{0.2pt}}
\linespread{1.08}
\frenchspacing\sloppy\sloppypar\raggedbottom

\title{\bfseries%
Equivariant convolution-based functions
on tensor fields}
\author{David W. Hogg, Ben Blum-Smith, and Soledad Villar}
\date{}

@@

\begin{definition}[vectors, scalars, and $k$-tensors] \label{def.tensors}
We say that $v\in \mathbb R^d$ is a vector (or 1-tensor) if $O(d)$ acts on $v$ via the canonical action. Namely, for any $g\in O(d)$ we have $g\cdot v = M(g)\, v$ where $M(g)\in \mathbb R^{d\times d}$ is the the matrix representation of $g$. We say that $a\in \mathbb R$ is a scalar (or 0-tensor) if $O(d)$ acts trivially on $a$. Namely $g\cdot a = a$.
We say that $T\in (\mathbb R^d)^{\otimes k}$ is a $k$-tensor if $O(d)$ acts on $T$ as follows:
$O(d)$ acts on a rank-1 $k$-tensor by $g\cdot (v_{1}\otimes\ldots \otimes v_k) = (g\cdot v_1)\otimes \ldots \otimes (g\cdot v_k)$ and it acts on higher rank tensors by linearly extending the actions to linear combinations of such rank-1 $k$-tensors.
\end{definition}

@@

\begin{definition}[pseudovectors, pseudoscalars, and $k$-pseudotensors]
We say that $B\in (\mathbb R^d)^{\otimes k}$ if a $k$-pseudotensor if it is acted by $O(d)$ by the following action: $g\odot B := \det(g)\, g \cdot B$ where $g\cdot B$ is the canonical action on $k$-tensors as in Definition \ref{def.tensors}. The pseudovectors and pseudoscalars are the 1-pseudotensors and 0-pseudotensors respectively.
\end{definition}

@@

\begin{definition}[permutations of tensor indices]
Given a $k$-tensor or $k$-pseudotensor $T\in (\mathbb R^d)^{\otimes k}$ and $\sigma\in S_k$ a permutation. We consider $T^\sigma$ the result of permuting the indices of $T$ according to $\sigma$:
\begin{equation}
    [T^\sigma]_{i_1, \ldots, i_k} := [T]_{\sigma^{-1}(i_1), \ldots, \sigma^{-1}(i_k)}
\end{equation}
\end{definition}

@@

\section{Expressive power of this architecture}
We can express all linear translation equivariant functions (by just using convolutions assuming filters are same size of the image).

We can cite that convolutions with small patches may be seen as low pass filters in some way. Reference?

\section{Construction of equivariant geometric convolution filters}
In the above we defined nonlinear equivariant geometric operators acting on $d$-dimensional images on the $d$-torus.
These, in turn, are constructed from products of the outputs of equivariant geometric convolution filters.
Here we show how to find all such filters by a group-averaging technique.
